PLAE enables quick and easy searching across one million ocular transcriptomes

### 2022-06-07 manuscript_00.Rmd
replace gene

diff_tab_big <- scEiaD_2020_v01 %>% tbl('diff_testing') %>% 
  filter(Group == 'CellType (Predict)', padj < 1e-5, (log2FoldChange) > 2) %>% 
  collect()

diff_tab_big2 <- diff_tab_big %>% 
  separate(Gene, into = c('Symbol', 'ID'), sep = ' ') %>% 
  mutate(ID = gsub('\\(|\\)', '', ID)) %>% 
  filter(Symbol == '-' | Symbol == 'NA')

@@

```{r}

spec_diff <- diff_tab_big %>% 
    filter(Gene %in% diff_tab$Gene) %>% # only retain diff genes across "All"
    group_by(Gene, Organism, Base) %>% 
    summarise(Count = n()) %>% 
    filter(Count > 1) %>% summarise(Count = n()) %>% filter(Count < 3) %>% mutate(go = paste(Gene, '___', Organism, sep = '')) %>% pull(go)

diff_tab_big %>% 
    filter(Gene %in% diff_tab$Gene) %>% # only retain diff genes across "All"
    group_by(Gene, Organism, Base) %>% 
    summarise(Count = n()) %>% 
    filter(Count > 1) %>% mutate(go = paste(Gene, '___', Organism, sep = '')) %>% filter(go %in% (spec_diff))  %>% select(-go) %>% ungroup() %>% group_by(Organism, Base) %>% slice_max(n = 10, order_by = Count) %>% filter(Base == 'Rod')
```

@@

```{r}
# consistently diff genes
consistent <- diff_tab %>% 
  group_by(Gene, Base)  %>% 
  summarise(Count = n()) %>% 
  filter(Count > 1) %>% 
  mutate(k = paste(Gene, Base, sep = '_'))

diff_tab %>% mutate(k = paste(Gene, Base, sep = '_')) %>% 
  filter(k %in% consistent$k) %>% 
  group_by(Base, Organism) %>% 
  summarise(Count = n()) %>% 
  rename(`CellType (Predict)` = Base, `Differentially Expressed Gene Count` = Count) %>% 
  pivot_wider(values_from = `Differentially Expressed Gene Count`, names_from = Organism) %>% replace(is.na(.),0) %>% 
  as_tibble() %>% 
  flextable()

@@

```{r}
# flip flop diff genes
down_tab <- scEiaD_2020_v01 %>% tbl('diff_testing') %>% 
  filter(Group == 'CellType (Predict)', padj < 1e-5, (log2FoldChange) < -2) %>% 
  collect() %>% 
  filter(Against == 'All', Base %in% (meta_filter %>% filter(Organ == 'Eye') %>% pull(CellType_predict) %>% unique())) %>% mutate(k = paste(Gene, Base, sep = '_'))

diff_tab %>% mutate(k = paste(Gene, Base, sep = '_')) %>% 
  filter(k %in% down_tab$k, Base %in% c('AC/HC Precursor', 'Amacrine Cell', 'Astrocyte', 'Bipolar Cell', 'Cone', 'Early RPC',' Horizontal Cell','Late RPC','Microglia','Muller Glia', 'Neurogenic Cell', 'Photoreceptor Precursor','Retinal Ganglion Cell','Rod','Rod Bipolar Cell','RPC','RPE' )) %>% 
  arrange(padj)

diff_tab_big %>% 
    filter(Gene %in% diff_tab$Gene) %>% # only retain diff genes across "All"
    group_by(Gene, Base, Organism) %>% 
    summarise(Count = n()) %>% 
    filter(Count > 8, Base == 'Rod') %>% arrange(-Count)
```

>>>>

diff_tab_big <- scEiaD_2020_v01 %>% tbl('diff_testing') %>% 
  filter(Group == 'CellType (Predict)', padj < 1e-5, (log2FoldChange) > 2) %>% 
  collect() %>% 
  separate(Gene, into = c('Symbol', 'ID'), sep = ' ') %>% 
  mutate(ID = gsub('\\(|\\)', '', ID))

diff_tab_big2 <- diff_tab_big %>% 
  filter(Symbol == '-' | Symbol == 'NA')

diff_tab_big2 <- 
  diff_tab_big2 %>% 
  left_join(hs_mm, by = c('ID' = 'GENEID')) %>% 
  select(-Symbol) %>% 
  rename(Symbol = SYMBOL) %>% 
  relocate(Symbol)

diff_tab_big <- bind_rows(
  diff_tab_big %>% filter(!Symbol %in% c('-', 'NA')),
  diff_tab_big2
)

@@

```{r}
spec_diff <- diff_tab_big %>% 
  filter(Symbol %in% diff_tab$Symbol) %>% # only retain diff genes across "All"
  group_by(Symbol, Organism, Base) %>% 
  summarise(Count = n()) %>% 
  filter(Count > 1) %>% summarise(Count = n()) %>% filter(Count < 3) %>% mutate(go = paste(Symbol, '___', Organism, sep = '')) %>% pull(go)

diff_tab_big %>% 
  filter(Symbol %in% diff_tab$Symbol) %>% # only retain diff genes across "All"
  group_by(Symbol, Organism, Base) %>% 
  summarise(Count = n()) %>% 
  filter(Count > 1) %>% mutate(go = paste(Symbol, '___', Organism, sep = '')) %>% 
  filter(go %in% (spec_diff))  %>% 
  select(-go) %>% 
  ungroup() %>% 
  group_by(Organism, Base) %>% 
  slice_max(n = 10, order_by = Count) %>% 
  filter(Base == 'Rod')
```

@@

```{r}
# consistently diff genes
consistent <- diff_tab %>% 
  group_by(Symbol, Base)  %>% 
  summarise(Count = n()) %>% 
  filter(Count > 1) %>% 
  mutate(k = paste(Symbol, Base, sep = '_'))

diff_tab %>% mutate(k = paste(Symbol, Base, sep = '_')) %>% 
  filter(k %in% consistent$k) %>% 
  group_by(Base, Organism) %>% 
  summarise(Count = n()) %>% 
  rename(`CellType (Predict)` = Base, `Differentially Expressed Gene Count` = Count) %>% 
  pivot_wider(values_from = `Differentially Expressed Gene Count`, names_from = Organism) %>% replace(is.na(.),0) %>% 
  as_tibble() %>% 
  flextable()

@@

```{r}
# flip flop diff genes
down_tab <- scEiaD_2020_v01 %>% tbl('diff_testing') %>% 
  filter(Group == 'CellType (Predict)', padj < 1e-5, (log2FoldChange) < -2) %>% 
  collect() %>% 
  filter(Against == 'All', Base %in% (meta_filter %>% filter(Organ == 'Eye') %>% pull(CellType_predict) %>% unique())) %>% mutate(k = paste(Symbol, Base, sep = '_'))

diff_tab %>% mutate(k = paste(Symbol, Base, sep = '_')) %>% 
  filter(k %in% down_tab$k, Base %in% c('AC/HC Precursor', 'Amacrine Cell', 'Astrocyte', 'Bipolar Cell', 'Cone', 'Early RPC',' Horizontal Cell','Late RPC','Microglia','Muller Glia', 'Neurogenic Cell', 'Photoreceptor Precursor','Retinal Ganglion Cell','Rod','Rod Bipolar Cell','RPC','RPE' )) %>% 
  arrange(padj)

diff_tab_big %>% 
  filter(Symbol %in% diff_tab$Symbol) %>% # only retain diff genes across "All"
  group_by(Symbol, Base, Organism) %>% 
  summarise(Count = n()) %>% 
  filter(Count > 8, Base == 'Rod') %>% arrange(-Count)
```
